---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->


```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```



```{r}
# load the library xlsx
library(xlsx)

# read the raw data
myXl <- read.xlsx("./inst/extdata/oilfield_100w_raw_data.xlsx", 
                  sheetIndex = 1, stringsAsFactors=FALSE)
```



## Misstyped data

```{r}
unique(myXl$Wellname)
```

## Pattern detection
```{r rows.print=15}
# using a template to find out which well names do not follow a pattern
myXl[!grepl("PSCO-[M,O,P,,Q,R,S][0-9][0-9][0-9]-[T,L,S]S", myXl$Wellname),]
```

The result is that we get the observations (rows) that have incorrect well names. They are 11 instances.
This is much better than visually inspecting them in a spreadsheet, isn't it?

What are the type of offences?

* Incorrect well number: PSCO-M0007-TS, PSCO-M0026-TS
* Platform omitted: PSCO-027-TS
* Platform in lowercase: PSCO-r015-LS, PSCO-m016-LS
* Incorrect field name: PiSCO-R009-SS, PISCO-R027-LS
* Incorrect completion type: PSCO-R022-T, PSCO-Q019-L, PSCO-Q001-S
* Extra spaces in the name: PSCO-S019 -LS


## Fix the well name

Some can be fixed manually and other can be done automatically with a script.
In our particular case we only have 100 wells but what about if we have 1000, or 5000? Doin,g it manually is not an option. Some are quickyly fixable some others are more challenging. Let's start by the easier ones.

Always go from the more general to the more particular.

```{r}
# lowercase to uppercase
  myXl$Wellname <- toupper(myXl$Wellname)

# show the wells with issues
  myXl[!grepl("PSCO-[M,O,P,,Q,R,S][0-9][0-9][0-9]-[T,L,S]S", myXl$Wellname),]
```


```{r}
# removing spaces
  myXl$Wellname <- gsub(" ", "", myXl$Wellname)

# show the wells with issues
  myXl[!grepl("PSCO-[M,O,P,,Q,R,S][0-9][0-9][0-9]-[T,L,S]S", myXl$Wellname),]
```


```{r}
# complete the completion type
# We have three completion type: SS, LS and TS

  myXl$Wellname <- gsub("-L$", "-LS", myXl$Wellname)    # the dollar sign at the end, means that 
  myXl$Wellname <- gsub("-S$", "-SS", myXl$Wellname)    # we are checking at the end of the string
  myXl$Wellname <- gsub("-T$", "-TS", myXl$Wellname)

# show the wells with issues
  myXl[!grepl("PSCO-[M,O,P,,Q,R,S][0-9][0-9][0-9]-[T,L,S]S", myXl$Wellname),]

```

Those were the easy ones. There are 5 more to go.

```{r}
# detect well names that are not named as PSCO-
  myXl[!grepl("^PSCO-", myXl$Wellname), ]        # the ^ character means look at the start of the string
```

### 

```{r}
# replace any characters before the first dash
  x <- myXl$Wellname

  pattern <- "^[^-]+"
  repl_with <- "PSCO"

  myXl$Wellname <- gsub(pattern, repl_with, x, perl = TRUE)

  # show the wells with issues
  myXl[!grepl("PSCO-[M,O,P,,Q,R,S][0-9][0-9][0-9]-[T,L,S]S", myXl$Wellname),]
```

Alright. We corrected the field name in the well name. 
There are still three more wells to go which problems are:

    PSCO-M0007-TS  long well number. It should be maximum of 3 digits
    PSCO-027-TS    missing platform number
    PSCO-M0026-TS  long well number. It should be maximum of 3 digits


### correcting longer well number (digits)

```{r}
# find wells with longer well number
x <- myXl$Wellname
myXl[grepl("[0-9]{4}-", x, perl=TRUE), ]     # 
```


```{r}
library(stringr)
x <- myXl$Wellname

allIndices <- 1:length(x)
incorrectIndices <- grep("[0-9]{4}-", x, perl=TRUE)
incorrectIndices

xt <- strsplit(x[incorrectIndices], "-")
fn <- unlist(lapply(xt, "[", 1))
xl <- unlist(lapply(xt, "[", 2))
ct <- unlist(lapply(xt, "[", 3))


first <- str_sub(xl, 1, 1)                 # extract the 1st character
xu <- unlist(str_extract_all(xl, "\\d+"))  # extract the numeric part only
xn <- as.numeric(xu) / 1000 * 1000
xp <- str_pad(xn, pad = "0", width = 3)
wn <- paste0(first, xp)
corrected <- paste(fn, wn, ct, sep = "-")

myXl$Wellname[incorrectIndices] <- corrected

  # show the wells with issues
  myXl[!grepl("PSCO-[M,O,P,,Q,R,S][0-9][0-9][0-9]-[T,L,S]S", myXl$Wellname),]

```

### correct the platform in the well name

```{r}
x <- myXl$Wellname
# pick up the well index
incorrectIndices <- !grepl("PSCO-[M,O,P,,Q,R,S][0-9][0-9][0-9]-[T,L,S]S", x)
# incorrectIndices

xt <- strsplit(x[incorrectIndices], "-")

fn <- unlist(lapply(xt, "[", 1))
xl <- unlist(lapply(xt, "[", 2))
ct <- unlist(lapply(xt, "[", 3))

xl 
platform <- myXl$Platform[incorrectIndices]

wn <- paste0(platform, xl)

corrected <- paste(fn, wn, ct, sep = "-")
corrected

myXl$Wellname[incorrectIndices] <- corrected

myXl[!grepl("PSCO-[M,O,P,,Q,R,S][0-9][0-9][0-9]-[T,L,S]S", myXl$Wellname),]

# finished the corrections of the well name
```



### Company

```{r}
myXl[!grepl("Oil Gains Co.", myXl$Company),]
# We don't get any return. All the company names are the same. Cool!
```

### Analyst

```{r}
summary(myXl$Analyst)
unique(myXl$Analyst)

```

```{r}
operators <- c("Aida", "Americo", "Andrew", "Camden", "Ibironke", "Kai", "Norman", 
               "Rodrigo", "Sam", "Tom", "Vivek")

# incorrect assignements
tocorrectIndices <- which(!myXl$Analyst %in% operators)
myXl[tocorrectIndices, ]
```

We can correct manually:

```{r}
myXl$Analyst[9] <- "Ibironke"
myXl$Analyst[c(31,32,32,33)] = "Rodrigo"
myXl$Analyst[c(78,83)] = "Andrew"
myXl$Analyst[c(37)] = "Tom"

# verify if we have incorrect assignements
  tocorrectIndices <- which(!myXl$Analyst %in% operators)
  myXl[tocorrectIndices, ]

```

There is only one observation left, the one with NA.
We will have to cross-reference it.
Let's find out who are the platform operators.

### Who is the platform operator?
In our particular case, who is the platform operator in "S"?
```{r}
# Who is the platform operator?
  table(myXl$Analyst, myXl$Platform)
```

For platform S, it could be Andrew or Camden.

```{r}
myXl[myXl$Analyst %in% c("Andrew", "Camden"), ]
```

There is not a real pattern here. We will assign the well to Andrew, make a note. Then we make a call to confirm.

```{r}
myXl$Analyst[c(98)] = "Andrew"

# verify if we have incorrect assignements
  tocorrectIndices <- which(!myXl$Analyst %in% operators)
  myXl[tocorrectIndices, ]

  # Analyst is complete  
```

## Check the field

```{r}
unique(myXl$Field)
# we have two additional keywords for the field that have to be corrected.
# let's find out what indices they have
```

```{r}
# verify for bad names in field
grep("[^PISCO]", myXl$Field)      # which means those which are not named like PISCO
```

```{r}
# which row index has NAs in it
myXl[which(is.na(myXl$Field)), ]
```


```{r}
# make the correction
myXl$Field[c(7,12,27)] <- "PISCO"

# verify for bad names in field
  grep("[^PISCO]", myXl$Field)    # which means those which are not named like PISCO
  
# which row index has NAs in it
myXl[which(is.na(myXl$Field)), ]  
  
# it has been fixed now  
```



## Add a column for the Completion Type
To close this chapter, let's add a c new variable (column) where we have only the Completion Type.
We can take advantage that the last two characters of the well name is the completion type.

    nchar(): number of character of the string

```{r}
substr(myXl$Wellname, nchar(myXl$Wellname)-1, nchar(myXl$Wellname))
```


```{r}
# assign the completion type to a new column
myXl$Completion <- substr(myXl$Wellname, nchar(myXl$Wellname)-1, nchar(myXl$Wellname))

```


## Location
```{r}
myXl$Location
```

```{r}
# we can test it by not following pattern or those which are NA
pattern <- "[MQRS][0-9]{3}"

# test that Location follows the pattern
grep(pattern, myXl$Location, invert = TRUE)
# there is one non-compliant index 
# which matches what we see above
```

```{r}
myXl$Location <- substr(myXl$Wellname[19], nchar(myXl$Wellname[19])-6, nchar(myXl$Wellname[19])-3)
```


```{r}
# test that Location follows the pattern
grep(pattern, myXl$Location, invert = TRUE)
# we are done here
```


## Platform

```{r}
# verify which indices do not comply for platform
grep("[MQRS]", myXl$Platform, invert = TRUE)
# only two not following
# since the well name is already corrected, let's use it
```

```{r}
myXl$Platform <- substr(myXl$Wellname, nchar(myXl$Wellname)-6, nchar(myXl$Wellname)-6)

# verify which indices do not comply for platform
grep("[MQRS]", myXl$Platform, invert = TRUE)
# we are done here
```


```{r}
# load the library xlsx
library(xlsx)

# read the raw data
myXl <- read.xlsx("./inst/extdata/oilfield_100w_raw_data.xlsx", 1)

# lowercase to uppercase
  myXl$Wellname <- toupper(myXl$Wellname)



# removing spaces
  myXl$Wellname <- gsub(" ", "", myXl$Wellname)

  
  

# complete the completion type
# We have three completion type: SS, LS and TS

  myXl$Wellname <- gsub("-L$", "-LS", myXl$Wellname)    # the dollar sign at the end, means that 
  myXl$Wellname <- gsub("-S$", "-SS", myXl$Wellname)    # we are checking at the end of the string
  myXl$Wellname <- gsub("-T$", "-TS", myXl$Wellname)

# show the wells with issues
  myXl[!grepl("PSCO-[M,O,P,,Q,R,S][0-9][0-9][0-9]-[T,L,S]S", myXl$Wellname),]
```


```{r}
# remove non compliant field names
  # myXl$Wellname <- gsub("(^[PSCO])", "PSCO", myXl$Wellname)

# show the wells with issues
  myXl[!grepl("PSCO-[M,O,P,,Q,R,S][0-9][0-9][0-9]-[T,L,S]S", myXl$Wellname),]
```



```{r}
# test
test <- c("PiSCO", "pISCO", "PSCO")

gsub("^(PSCO)", "PSCO", test)

```



```{r}
# show wells that do have incorrect field name
myXl[!grepl("^PSCO-", myXl$Wellname), ]
```

```{r}
# same but saving the data frame
df <- myXl
pattern <- "^(?!PSCO)"
myXl[grepl(pattern, myXl$Wellname, perl = TRUE), ]
```


```{r}
# attempt to replace PISCO. nOT WORKING
df <- myXl
pattern <- "(?!PSCO+)"
df$Wellname <- gsub(pattern, "\\1", df$Wellname, perl = TRUE)

df[!grepl("PSCO-[M,O,P,,Q,R,S][0-9][0-9][0-9]-[T,L,S]S", df$Wellname),]
```


```{r}
# show wellls that do not comply with well naming pattern
df[!grepl("PSCO-[M,O,P,Q,R,S][0-9][0-9][0-9]-[T,L,S]S", df$Wellname),]
```

