---
title: "5.1 Data Introspection"
output:
  pdf_document:
    toc: yes
  html_document:
    keep_md: yes
    toc: yes
---


<!-- README.md is generated from README.Rmd. Please edit that file -->


```{r setup, include=F, error=T, message=F, warning=F}
knitr::opts_chunk$set(echo=T, comment="#:>", error=T, warning=F, message = F, fig.align = 'center', collapse=TRUE)
```


This section is about getting familiar with our data. We will be using functions to know the size of our table or data frame, the names of the columns or variables, the staructure of the data and the type of data for each of the variables or colummns.

## Read the `raw data` again

```{r}
# load the library xlsx
library(xlsx)

# read the raw data
myXl <- read.xlsx("../extdata/oilfield_100w_raw_data.xlsx", 
                  sheetIndex = 1, stringsAsFactors = FALSE)

```

## Printing the `head` of the dataframe

Let's print 6 rows of data with the function head(). You will see a long printing. We will fix this in a minute. Read on.

```{r rows.print=10}
# the function head() prints the first 6 rows
# to print the last 6 rows use tail()
print(head(myXl))
```


It looks pretty long.
Let's try with a package that adds better printing capabilities: `tibble`.

## Install `tibble`

Install it with `install.packages("tibble")`

```{r}
library(tibble)      # load the package
myXl <- as_tibble(myXl)      # convert the data frame to a tibble
```

```{r}
head(myXl)
```


```{r}
tail(myXl)
```


Now it looks much better.

## dimensions of the data frame: `dim`
Let's use some R functions to find out more about our data.

```{r}
# get the dimensions of the table.
dim(myXl)
```

Our table has `r dim(myXl)[1]` rows and `r dim(myXl)[2]` columns.

## names of the columns: `names`
These are the names of the variables or columns:

```{r}
names(myXl)
```

## summary of the data: `summary`

```{r}
# A summary of all the variables.
# Notice the difference between numerical and non-numerical variables
summary(myXl)
```


## structure of the data: `str`

```{r}
# show the data type structure of the table
str(myXl)
```

```{r}
# outr table is one of R data structures along with vectors, matrices, arrays and lists.
class(myXl)
```


## data types: `typeof`
Let's find out what type of variable we've got in our table:

```{r}
typeof(myXl$Wellname)
typeof(myXl$Fluid)
typeof(myXl$IPR_RESTEMP)
```

## using `sapply`, `length`, `sort`

We can do all the column names in one shot with `sapply`.

```{r}
dataTypes <- sapply(myXl, typeof, simplify = "array")
typeof(dataTypes)
length(dataTypes)
sort(dataTypes)

```

## An inventory of the kind of data we have: `table`

```{r}
table(dataTypes)
```



