---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->


```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```



```{r}
# load the library xlsx
library(xlsx)

# read the raw data
myXl <- read.xlsx("../extdata/oilfield_100w_raw_data.xlsx", 
                  sheetIndex = 1, stringsAsFactors=FALSE)

```



## Data introspection
Let's print 6 rows of data with the function head(). You will see a long printing. We will fix this in a minute. Read on.

```{r rows.print=25}
# the function head() prints the first 6 rows
# to print the last 6 rows use tail()
print(head(myXl))
```


It looks pretty long.
Let's try with a package that adds better printing capabilities: `tibble`.

## Install `tibble`

Install it with `install.packages("tibble")`

```{r}
library(tibble)      # load the package
as_tibble(myXl)      # convert the data frame to a tibble
```

Now it looks much better.

### dimensions of the data frame: `dim`
Let's use some R functions to find out more about our data.

```{r}
# get the dimensions of the table.
dim(myXl)
```

Our table has `r dim(myXl)[1]` rows and `r dim(myXl)[2]` columns.

### names of the columns: `names`
These are the names of the variables or columns:

```{r}
names(myXl)
```

### summary of the data: `summary`

```{r}
# A summary of all the variables.
# Notice the difference between numerical and non-numerical variables
summary(myXl)
```


### structure of the data: `str`

```{r}
# show the data type structure of the table
str(myXl)
```

```{r}
# outr table is one of R data structures along with vectors, matrices, arrays and lists.
class(myXl)
```


### data types: `typeof`
Let's find out what type of variable we've got in our table:

```{r}
typeof(myXl$Wellname)
typeof(myXl$Fluid)
typeof(myXl$IPR_RESTEMP)
```

### using `sapply`, `length`, `sort`

We can do all the column names in one shot with `sapply`.

```{r}
dataTypes <- sapply(myXl, typeof, simplify = "array")
typeof(dataTypes)
length(dataTypes)
sort(dataTypes)

```

### An inventory of the kind of data we have: `table`

```{r}
table(dataTypes)
```


